---
title: "Assorted Musings markdown"
output: html_document
date: "2024-04-18"
---

THis R Makrdown is mostly just for me to muse and play with data while recording results in a somewhat useful way.
I expect this document to get really long a tedious.  WHen a that happens I will switch to a new one.   Eventually once I find a analysis schema that I like I will move this all
to a R script so it can run a bit more smoothely and doesnt need all the musing.  Until then this is what we got.

Note:  I will do my best to create libraries/functions where helpful and needed.   I will also try and explain the goal and intent of what I want to do before each code block
Unsure if this will actually be helpful and it probably wont be in the proper YAML code.  Sorryyyyyy

```{r setup, include=FALSE}
library(lubridate, warn.conflicts = FALSE)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
knitr::opts_chunk$set(echo = TRUE)

```


First we must import all the data from Piet's long term data.  Some important notes
1. Piets data does not have that much information on amphibians themselves.   Or at least not adults   Has really good young populations and other data but if you want data on the adult frogs including load we must look at Cherie's and Caitlin's data.  These are more complete yet cover fewer ponds.

```{r}
survey_df = read.csv("/Users/canonmallory/Library/Mobile Documents/com~apple~CloudDocs/Documents/School/UCSB/Briggs Lab/Climate Grant/Raw_Data_East_Bay_Work_vs_Load/SURVEY_SPP_piet4.csv")
site_info_df = read.csv("/Users/canonmallory/Library/Mobile Documents/com~apple~CloudDocs/Documents/School/UCSB/Briggs Lab/Climate Grant/Raw_Data_East_Bay_Work_vs_Load/Site_Info_piet6.csv")
water_qual_df = read.csv("/Users/canonmallory/Library/Mobile Documents/com~apple~CloudDocs/Documents/School/UCSB/Briggs Lab/Climate Grant/Raw_Data_East_Bay_Work_vs_Load/Water_Quality_piet5.csv")
wetland_info_df = read.csv("/Users/canonmallory/Library/Mobile Documents/com~apple~CloudDocs/Documents/School/UCSB/Briggs Lab/Climate Grant/Raw_Data_East_Bay_Work_vs_Load/Wetland_Info_piet9.csv")
```


Lets make a histogram to look at water quality information really quick.   I am most interested in 
A. Temp
B. Salinity
C. pH

But lets add the other ones too.  
Conductivity
TDS (total dissolved solids)
DO_Pct (unsire what this is)(i should google that one)
DO_mg (also uknown....  should google stuff)
Secchi

```{r}
years = unique(water_qual_df$Year)
first_year = min(years)
last_year = max(years)

temp_hist = ggplot(water_qual_df, aes(x = WaterTemp_C))+
  geom_histogram(bins = 100)

conductivity_hist = ggplot(water_qual_df, aes(x = Conductivity_uS_cm)) +
  geom_histogram(bins = 100)

ph_hist = ggplot(water_qual_df, aes(x = pH)) +
  geom_histogram(bins = 100)

secchi_hist = ggplot(water_qual_df, aes(x = SecchiDepth_cm)) +
  geom_histogram(bins = 100)

TDS_hist = ggplot(water_qual_df, aes(x = TDS_g_l)) +
  geom_histogram(bins = 100)

water_quality_all_years_fig = ggarrange(temp_hist, conductivity_hist, ph_hist, secchi_hist, TDS_hist)
annotate_figure(water_quality_all_years_fig,
                top = text_grob(paste("Water Quality (", toString(first_year), "-", toString(last_year), ")", sep=""))
)
```

Okay now lets compair each of the water quality variables to the results we see during drought years.
We should also addin the flooding years but I don't have those redily available so that is a 
TODO:  ADD FLOODING YEARS

For now we will consider drought years to be 2020-2022.

```{r}

drought_years = c(2020,2021,2022)
drought_years_df <- water_qual_df[water_qual_df$Year %in% drought_years, ] #Gets all the years that are in our designated drought years

#Creates fill titles.  WIll be consistant across all the histograms
all_years_fill = paste(toString(first_year), "-", toString(last_year), sep="")
drought_yeasr_fill = paste(toString(min(drought_years)), "-", toString(max(drought_years)), sep="")

temp_hist = ggplot()+
  geom_histogram(aes(x = water_qual_df$WaterTemp_C, fill=all_years_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(water_qual_df$WaterTemp_C)), fill=all_years_fill))+
  geom_histogram(aes(x = drought_years_df$WaterTemp_C, fill=drought_yeasr_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(drought_years_df$WaterTemp_C)), fill=all_years_fill)) +
  theme(legend.position = "none")


conductivity_hist = ggplot()+
  geom_histogram(aes(x = water_qual_df$Conductivity_uS_cm, fill=all_years_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(water_qual_df$Conductivity_uS_cm)), fill=all_years_fill))+
  geom_histogram(aes(x = drought_years_df$Conductivity_uS_cm, fill=drought_yeasr_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(drought_years_df$Conductivity_uS_cm)), fill=all_years_fill)) +
  xlim(0, 2500)+
  theme(legend.position = "none")


ph_hist = ggplot()+
  geom_histogram(aes(x = water_qual_df$pH, fill=all_years_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(water_qual_df$pH)), fill=all_years_fill))+
  geom_histogram(aes(x = drought_years_df$pH, fill=drought_yeasr_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(drought_years_df$pH)), fill=all_years_fill)) +
  theme(legend.position = "none")


secchi_hist = ggplot()+
  geom_histogram(aes(x = water_qual_df$SecchiDepth_cm, fill=all_years_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(water_qual_df$SecchiDepth_cm)), fill=all_years_fill))+
  geom_histogram(aes(x = drought_years_df$SecchiDepth_cm, fill=drought_yeasr_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(drought_years_df$SecchiDepth_cm)), fill=all_years_fill)) +
  theme(legend.position = "none") + 
  xlim(0, 200) + 
  ylim(0, 100)

TDS_hist = ggplot()+
  geom_histogram(aes(x = water_qual_df$TDS_g_l, fill=all_years_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(water_qual_df$TDS_g_l)), fill=all_years_fill))+
  geom_histogram(aes(x = drought_years_df$TDS_g_l, fill=drought_yeasr_fill), bins = 100, alpha = 0.5)+
  geom_vline(aes(xintercept = mean(na.omit(drought_years_df$TDS_g_l)), fill=all_years_fill)) +
  theme(legend.position = "none") +
  xlim(0, 2000)


water_quality_all_years_fig = ggarrange(temp_hist, conductivity_hist, ph_hist, secchi_hist, TDS_hist, common.legend = TRUE)
annotate_figure(water_quality_all_years_fig, top = text_grob("Water Quality Drought vs All Years"))
```

